---
title: "tisB-istR analysis"
format: gfm
editor: visual
---

# Analysis of tisB-istR genes

Following the analysis shown in `selene_stats.md`, I'm doing the same analysis with a different set of genes.

I'll skip the parametric models and go straight to the ANOVA with Monte Carlo that I described previously.

```{r setup, message=FALSE}
library(tidyverse)
theme_set(theme_classic())
source("R/recreate_raw_data.R")


tisB_region <- read_gene_freq_data("data/tisB_istR_region.csv") 
tisB_path <- read_gene_freq_data("data/tisB_istR_path.csv") 


crap.loop.time <- system.time({ # takes about a minute on my machine
  raw_tisB_region_data <- recreate_raw(tisB_region) #%>%
  #arrange(category, gene.count)# This is running really slowly,
# Issues a message about data.table, whcih I thought was not loaded. Hmmm....
  raw_tisB_path_data <- recreate_raw(tisB_path)# %>%
  #arrange(category, gene.count)
})


p_region_tisB <- raw_plot(tisB_region)
p_path_tisB <- raw_plot(tisB_path)

# Really a bar graph will do fine here
p_tisB_region_b <- barplot_2(tisB_region) + 
  scale_fill_manual(name="gene number", values = c("#517c96", "#8D2048"))
p_tisB_path_b <- barplot_2(tisB_path) +
   scale_fill_manual(name="gene number", values = c("#517c96", "#8D2048"))


cowplot::plot_grid(p_tisB_path_b, p_tisB_region_b)
```
Now we just need to do the analysis of whetehr those distributions are different among categories.

```{r create_linear_models}
tisB_path_model <- lm(gene.count ~ category, data=raw_tisB_path_data)
summary(tisB_path_model)
```
```{r}
tisB_region_model <- lm(gene.count ~ category, data=raw_tisB_region_data)
summary(tisB_region_model)
```

```{r monte_carlo}
n <- 10000
#nrow.reg <- nrow(tisB_region)
#nrow.path <- nrow(tisB_path)
reg.f.vec <- vector("double", n)
path.f.vec <- vector("double", n)

set.seed(3230)
region.loop.time <- system.time({
  for(i in 1:n) {
  reg.f.vec[i] <- shuf_calc_f(raw_tisB_region_data)
}
})
set.seed(2112)
  for(i in 1:n) {
  path.f.vec[i] <- shuf_calc_f(raw_tisB_path_data)
  }

# Put simulated f vals in a data frame, and pull out observed f values from the model
f_vals <- data.frame(reg.sim.f = reg.f.vec,
                     path.sim.f = path.f.vec)
reg.f.real <- summary(aov(tisB_region_model))[[1]][1,4]
path.f.real <- summary(aov(tisB_path_model))[[1]][1,4]

```

## f values by region
```{r plot_region_histogram}
p_reg_hist <- ggplot(f_vals, aes(x=log10(reg.sim.f))) + 
  geom_histogram(bins = 100) + 
  geom_vline(xintercept = log10(reg.f.real), color="red")  + 
  ggtitle("data by region")
print(p_reg_hist)
```
The distribution of f values here is pretty weird: note the log scale. It is being caused by the fact that there aren't many ways to permute the labels on the data we have, so we end up with `r length(unique(f_vals$reg.sim.f))` unique f vals (despite `r n` simulations). Many of these have odd combinations: for instance if each region gets *either* 0 gene copies or 2 gene copies, but not a mix of both, the f value will be extreme. 

So I'm not very satisfied by this simulation. However, our observed f value is greater than `r round(100-100*sum(f_vals$reg.sim.f > reg.f.real)/nrow(f_vals), digits=2)`, indicating a p value of `r round(sum(f_vals$reg.sim.f > reg.f.real)/nrow(f_vals), digits=4)`.

## f values by pathotype

```{r plot_pathotype_histogram, warning=FALSE}
p_path_hist <- ggplot(f_vals, aes(x=path.sim.f)) + 
  geom_histogram(bins = 100) + 
  geom_vline(xintercept = path.f.real, color="red") + 
  ggtitle("data by pathology")
print(p_path_hist)
```


```{r}
f_clean <- f_vals%>%
  filter(is.finite(path.sim.f),
         path.sim.f > 1e-10,
         path.sim.f < 1e10)


p_path_hist_clean <- ggplot(f_vals %>%
                              filter(is.finite(path.sim.f),
                                     path.sim.f > 1e-10,
                                     path.sim.f < 1e10), 
                            aes(x=path.sim.f)) + 
  geom_histogram(bins = 100) + 
  geom_vline(xintercept = path.f.real, color="red") + 
  ggtitle("data by pathology")
print(p_path_hist_clean)
```





